Evaluating the Performance and Energy Efficiency of the \textsc{COSMO-ART} model system

### 2014-04-25 paper_enahpc14/Sec3:Environment-setup.tex
Manuel's changes

\begin{itemize}
\item \monch is a 10-rack NEC-provided clus\-ter at CSCS composed
  of 312 compute  nodes. %and utilised by members  of the Swiss Platform
  %for Advanced Scientific  Computing (PASC) project~\citep{PASC}
  Each node is comprised of two Intel Ivy Bridge EP  E5-2660 ten-core processors
  operating at  2.2 GHz, equipped with 32GB of DDR3 1600\,MHz  RAM and connected  via InfiniBand Mellanox  SX6036 and FDR switches. For our  experiments, 
  a full rack constituted of 52 nodes was used. \monch is  slated to stay
  in  service  without  hardware  upgrade  for  the  duration  of  the
  Exa2Green  project,  allowing   an  identical  configuration  to  be
  recreated for future assessments of the baseline.\\

\item \pilat is a dual-socket eight-core Intel Sandy
  Bridge EP based cluster at  CSCS composed of 42 compute nodes. %and used as Piz Daint pre-post  processing cluster 
  The 2 login and
  42 compute nodes are consolidated into 11 twin-pair Intel E5-Series DALCO
  r2264i4t 2U scalable compute modules, each containing 4 compute
  nodes based  on two Intel Xeon E5-2670 processors operating  at 2.6\,GHz equipped with 64GB of DDR3 1600\,MHz RAM and connected  via InfiniBand Mellanox SX6036 and FDR switches. 
  For our  experiments, a  full rack constituted  of 42 standard  compute nodes
  was used. \pilat  is based  on  Intel's second-generation
  processors, conventional in  HPC systems and known to  be more power
  consuming than its Ivy Bridge successor.\\

\item \tinto is  a heterogeneous cluster composed of
  28 compute nodes at UJI. For our  experiments only a subset of 16 homogeneous nodes was considered. Each of them is composed dual-socket Intel Xeon E5645 hexa-core processors 
  running at 2.40\,GHz, equipped with 24\,GB  of DDR3 1333\,MHz and connected via Infiniband QDR with a Mellanox MTS\-3600 switch.

@@

We present here the two different  frameworks deployed on the testbed HPC systems to
measure  performance and  power  consumption of the baseline execution of \cosmoart.

@@

For our experiments at Jaume I University (UJI), we employed a version
of the integrated framework presented in~\cite{energy13} that employs the \pmlib library in
combination with Extrae and Paraver, which  are profiling/tracing and
visualisation tools, respectively.

\cosmoart  is compiled and linked with the Extrae instrumentation library,
which automatically  instruments the Fortran~90 code of the model, e.g., to record MPI calls.  Next, \cosmoart  is  run on  the  nodes,  thus dissipating  certain
amount  of power.   These  nodes are connected to power distribution units (PDUs)
APC 8653 that account for the dissipated power with a sampling frequency of 1\,Hz. The power data captured by these PDUs is collected into the tracing server. The client,  meanwhile, sends a start/stop  primitive  in  order  to  gather  captured  data  by  the PDUs.

Once  the run  is finished,  a power-performance trace file containing the power 
data received from the tracing server and the instrumentation post-process is generated.
This file is finally loaded in Paraver in order to visualise the performance trace and to correlate it with the power profile generated by \cosmoart.

@@

A snapshot of the code,  which includes, at least conceptually, all the
information needed  to reproduce the  energy-to-solution benchmarks of
\cosmoart, was produced and run  on a 1040 cores using 20~MPI
tasks  per node on  \monch and  on a  1344 cores  using 16~MPI
tasks per node on  \pilat.  The calculated region was mapped
to the participating processors using a 2D-partitioning strategy.  The
distribution along the $x$ and $y$ coordinates was defined by setting:
$nprocx=40$  and $nprocy=26$  for \monch  and  $nprocx=28$ and
$nprocy=24$  as  $nprocx$  is   usually  kept  bigger  than  $nprocy$.
Besides,  as this  version doesn't  make use  of the  traditional GRIB
library, we  specified $nprocio=0$  for GRIB I/O.   Hyperthreading was
not considered in this study  as it previously revealed that it always
led to higher energy-to-solution.

>>>>

\begin{itemize}
\item \monch is a 10-rack NEC-provided clus\-ter at CSCS composed
  of 312 compute  nodes. %and utilised by members  of the Swiss Platform
  %for Advanced Scientific  Computing (PASC) project~\citep{PASC}
  Each node is comprised of two Intel Ivy Bridge EP  E5-2660 ten-core processors
  operating at  2.2 GHz, equipped with 32GB of DDR3 1600\,MHz  RAM and connected  via InfiniBand Mellanox  SX6036 and FDR switches. For our  experiments, 
  a full rack constituted of 52 nodes was used. % \monch is  slated to stay
  %in  service  without  hardware  upgrade  for  the  duration  of  the
  %Exa2Green  project,  allowing   an  identical  configuration  to  be
  %recreated for future assessments of the baseline.\\

\item \pilat is a dual-socket eight-core Intel Sandy
  Bridge EP based cluster at  CSCS composed of 42 compute nodes. %and used as Piz Daint pre-post  processing cluster 
  %The 2 login and 42 compute nodes %are consolidated into 11 twin-pair Intel E5-Series DALCO
  %r2264i4t 2U scalable compute modules, each containing 4 compute
  %nodes based 
  %consisting of 
  Each node is comprises two Intel Xeon E5-2670 processors operating  at 2.6\,GHz equipped with 64GB of DDR3 1600\,MHz RAM and connected  via InfiniBand Mellanox SX6036 and FDR switches. 
  For our  experiments, a full rack constituted of 42 compute nodes
  was used. \pilat  is based  on  Intel's second-generation
  processors, conventional in  HPC systems and known to  be more power
  consuming than its Ivy Bridge successor.\\

\item \tinto is  a heterogeneous cluster composed of
  28 compute nodes at UJI. For our  experiments only a subset of 16 homogeneous nodes was considered. Each node is composed of two Intel Xeon E5645 hexa-core processors 
  running at 2.4\,GHz, equipped with 24\,GB  of DDR3 1333\,MHz and connected via Infiniband QDR with a Mellanox MTS\-3600 switch.

@@

We present here the two different  frameworks deployed on the testbed HPC systems to
measure  performance, power  dissipation and energy consumption of the baseline execution of \cosmoart.

@@

For experiments at UJI on \tinto, we employed a version
of the integrated framework presented in~\cite{energy13} that employs the \pmlib library in
combination with Extrae and Paraver, which  are profiling/tracing and
visualisation tools, respectively.

To employ our framework, \cosmoart is first compiled and linked with the Extrae instrumentation library, which automatically  instruments the Fortran~90 code of the model, e.g., to record MPI calls.  Next, \cosmoart  is  run on  the  nodes,  thus dissipating  certain amount  of power. The  nodes are connected to power distribution units (PDUs)
APC 8653 that account for the dissipated power with a sampling rate of 1\,Hz. The power data captured by these PDUs is collected by a \pmlib tracing server running on a separate machine. The client, running on the target nodes, sends start/stop  primitive in order to collect
the power data during the \cosmoart run. Once  the run  is finished,  a power-performance trace file containing the power data is received from the tracing server and the instrumentation post-process is generated. A merged trace file is finally loaded in Paraver to visualise and correlate the performance trace and with the power profile generated by \cosmoart. This procedure allows to detect power bottlenecks, basically constituted by inefficient waiting methods.

@@

A snapshot of the code,  which includes, at least conceptually, all the
information needed  to reproduce the  energy-to-solution benchmarks of
\cosmoart, was produced and run on 1040 cores using 20~MPI
tasks  per node on  \monch, on 1344 cores  using 16~MPI
tasks per node on  \pilat and on 192 cores  using 12~MPI
tasks per node on \tinto.  The calculated region was mapped
to the participating processors using a 2D-partitioning strategy.  The
distribution along the $x$ and $y$ coordinates was defined accordingly, keeping in mind
% $by setting:
%$nprocx=40$  and $nprocy=26$  for \monch  and  $nprocx=28$ and
%$nprocy=24$  as  
to keep $x$  bigger  than  $nprocy$.
Besides,  as this  version does not  make use  of the  traditional GRIB
library, we  specified $nprocio=0$  for GRIB I/O.   Hyperthreading was
not considered in this study  as it previously revealed that it always
led to higher energy-to-solution.

### 2014-04-25 paper_enahpc14/Sec4:Experimental-results.tex
Manuel's changes

In this section, we present our measurement methodology along with the
metrics considered and discuss energy efficiency and power-performance
traces of the model system.

@@

For the experiments with \tinto platform, we also analyse the contribution of the MPI library to the energy consumption and how the use of the blocking versus polling message-passing policies can potentially render energy savings. Specifically, the OpenMPI 1.6.5 library, 
features two operation modes, blocking and polling, which can be selected before \cosmoart
is launched. In the polling configuration (the detault mode) ``MPI'' continually polls
the network interface to check for the completion of an event (e.g. send or receive).
This usually yields low latency but high CPU utilization. Thus, one can expcet that this mode 
attains the best performance, possibly at the cost a higher energy usage. In the blocking
mode, the CPU is yield to other processes/threads if there are no incoming messages.

@@

In  Figure~\ref{fig:3}, we  compare both  TTS (right  Y-axis)  and ETS
(left  Y-axis)  metrics  on  both  platforms.  As  expected,  Xeon  E5
outperforms Ivy Bridge EP, being  roughly 1.3x faster.  The reason for
that is  two-fold: \emph{i}) it has  higher clock frequency  than Ivy Bridge
(2.6\,GHz  against  2.2\,GHz),  and  \emph{ii})  it  aims at  computing  speed
regardless to  energy consumption.  In our experiments,  Ivy Bridge EP
showed the best energy-to-solution, reducing the energy consumption of
Xeon E5 by approximately 7\,\%.

The TTS and ETS results for a full-day simulation on \tinto are shown in Figure~\ref{fig:4}.  
As expected, the mean energy consumed by COSMO-ART when using the blocking MPI policy is about 
2\,W less than that consumed by the polling mode. This is due to the blocking-waits performed 
by the CPU cores while using the blocking mode, in contrast of the polling-waits that take 
place when polling mode is used. From the point of view of the execution time, there is a 
slight reduction when the blocking MPI policy is leveraged. One can expect that the costs of 
the interruptions used in the blocking MPI policy can lead to an increase of the total 
execution time, however, applications with extensive computing may show better perforamnce with 
the blocking modes. This plays in favour for our results since the energy consumed is also less 
by just changing the behavior of the OpenMPI library from the polling to the blocking policy.

@@

\begin{figure}[htbf]
  \includegraphics[width=0.5\textwidth]{Figs/Time_E2S_COSMO-ART-5.eps}
  \caption{Mean time-to-solution and energy-to-solution on
    \textsc{Tintorrum} for a 24h simulation using the polling and blocking MPI policies. The 95\% 
            confidence intervals of Student's $t$ distribution (10 samples) clearly indicates 
            an improvement for blocking policy in both TTS and ETS.}
  \label{fig:4}
\end{figure}

@@

A full tracing experiment is conducted on \tinto in order to capture an overall power
profile at a much finer resolution using the attached PDUs in each node. First we run \cosmoart 
for a full-day simulation and analyse the power consumption. Second we correlate performance 
with power traces of two 1-hor time frames during the day: mid-day (from 11h-12h) and
midnight (23h-24h). All the experiments were performed using 192 \cosmoart processes
mapping 12 processes in each of node.

% For all the experiments we also analyse the contribution of the MPI library to the energy 
% consumption and how the use of the blocking versus polling message-passing policies can 
% potentially render energy savings. Specifically, the OpenMPI 1.6.5 library installed in \tinto, 
% features two operation modes, blocking and modes, which can be selected before \cosmoart is 
% launched. In the polling configuration (the detault mode) ``MPI'' continually polls
% the netwrok interface to check for the completion of an event (e.g. send or receive).
% This usually yields low latency but high CPU utilization. Thus, can one expcet that this mode attanins the best performance, possibley at the cost a higher energy usage. In the blocking
% mode, the CPU is yield to other processes/threads if there are no incoming messages.

In order to obtain an initial overview of the power dissipation of the system model, we have 
obtained the power profiles using our \pmlib framework of a full-day simulation of \cosmoart 
leveraging the polling and blocking MPI policies, these results are depicted in the top/bottom 
plots of Figure~\ref{fig:9}, respectvely. The first observation for 
the polling MPI policy is the quite plain power profile it generates. However, the first and 
last hours of the full-day simulation dissipate slightly less power than the mid-day hours. We 
relate these variations due to the chemical reactions and aerosols, computed by ART, that are 
taking place during the daylight hours but not during night. In this case power consumption
 is about 3821\,W. Looking at the power profile using the blocking MPI policy, we 
can easily observe that the pattern is repeated favoring the reduction of power during 
nighttime hours in a more  noticeable way. The peak power dissipation in this mode is 3768\,W, 
which means a reduction of 50\,W in total or, in other words, 1\,\% of reductuon in the peak 
power. Also, we observe power  drops in each hour simulated, basically related to periods in 
which the processes were performing blocking-waits instead of polling-waits, thus yielding the 
CPU to a low usage during this phases. Our final observation is the reduction in execution time 
when the blocking MPI policy is leveraged in contrast with the polling MPI. One can expect that 
the costs of the interruptions used in the blocking can lead to an increase of the total 
execution time, however, applications with extensive computing may show better perforamnce with 
the blocking mode. This plays in favour for our results since the energy consumed is also less 
by just changing the behavior of the OpenMPI library from the polling to the blocking policy.

To gain more insights about the power-performance behavior of \cosmoart, we also have conducted 
experiments that show power-performance traces of 1 hour simulation. We limited the simulation 
time to this bound since the pattern of the computation performed by \cosmoart is repeated each 
hour, also because the weight of a full-day simulation trace can not be easily handled using 
our machines and memory available (about 45\,GB per trace file). To soften the problem we 
shrink the simulations to 1 hour during mid-day (11h-12h) and midnight (23h-24h), in which 
\cosmoart perform different internal computations, as shown in Figure~\ref{fig:9}. At the same
time we also leverage the two stated MPI policies: polling and waiting modes.

@@

\begin{figure*}[htbf]
  \centering
  \hspace{0.8cm}
  \scalebox{0.5}{\input{Figs/24hour.pstex_t}}
  \caption{24 hours simulation trace using the MPI blocking mode on \tinto.}
  \label{fig:9}
\end{figure*}

\begin{figure*}[htbf]
  \centering
  \scriptsize
  Simulation of 1 hour during midday (from 11h to 12h) leveraging the polling MPI policy\\
  \scalebox{0.5}{\input{Figs/11_12_blq0_figure.pstex_t}}\\
  Simulation of 1 hour during midday (from 11h to 12h) leveraging the blocking MPI policy\\
  \scalebox{0.5}{\input{Figs/11_12_blq1_figure.pstex_t}}\\
  1 hour simulation during midnight (from 23h to 24h) leveraging the polling MPI policy\\
  \scalebox{0.5}{\input{Figs/23_24_blq0_figure.pstex_t}}\\
  1 hour simulation during midnight (from 23h to 24h) leveraging the blocking MPI policy\\
  \scalebox{0.5}{\input{Figs/23_24_blq1_figure.pstex_t}}\\
  \begin{tabular}{rlp{-0.2cm}rlp{-0.2cm}rlp{-0.2cm}rlp{-0.2cm}rlp{-0.2cm}rl}
& &  & &  & &  & &  \\[-0.15cm]
Computation     & \multicolumn{1}{>{\columncolor[RGB]{  0,  , 255}}p{0.4cm}}{}  & & 
Waiting a msg.  & \multicolumn{1}{>{\columncolor[RGB]{255,  0  ,0}}p{0.4cm}}{}  & & 
Block. send     & \multicolumn{1}{>{\columncolor[RGB]{255,  0,174}}p{0.4cm}}{}  & & 
Synchronization & \multicolumn{1}{>{\columncolor[RGB]{179,  0,  0}}p{0.4cm}}{}  & & 
Wait/Wait all   & \multicolumn{1}{>{\columncolor[RGB]{235,  0,  0}}p{0.4cm}}{}  \\
& &  & &  & &  & &  \\[-0.15cm]
Immediate recv. & \multicolumn{1}{>{\columncolor[RGB]{100,100,177}}p{0.4cm}}{}  & &
I/O             & \multicolumn{1}{>{\columncolor[RGB]{172,174, 41}}p{0.4cm}}{}  & &
Group comm.     & \multicolumn{1}{>{\columncolor[RGB]{255,144, 26}}p{0.4cm}}{}  & &
Others          & \multicolumn{1}{>{\columncolor[RGB]{192,224,  0}}p{0.4cm}}{}  & &
Idle            & \multicolumn{1}{>{\columncolor[RGB]{117,195,255}}p{0.4cm}}{}  \\
\end{tabular}
  \caption{Power-perfomance traces during midday and midnight leveraging the polling and blocking MPI policies on \tinto.}
  \label{fig:11}
\end{figure*}

>>>>

In this section, we present our measurement methodology along with the
metrics considered and discuss about the performance-power-energy trade-off of the model system on the different platforms stated.

@@

For the experiments with \tinto platform, we also analyze the contribution of the MPI library to the energy consumption and how the use of the blocking versus polling message-passing policies can potentially render energy savings. Specifically, the OpenMPI 1.6.5 library, 
features two operation modes, blocking and polling, which can be selected before \cosmoart
is launched. In the polling configuration (the default mode) ``MPI'' continually polls
the network interface to check for the completion of an event (e.g. send or receive).
This usually yields low latency but high CPU utilization. Thus, one can expect that this mode 
attains the best performance, possibly at the cost a higher energy usage. In the blocking
mode, the CPU is yield to other processes/threads if there are no incoming messages.

@@

In  Figure~\ref{fig:3}, we  compare both  TTS (right  Y-axis)  and ETS
(left  Y-axis)  metrics  on  both  platforms.  As  expected,  Xeon  E5
outperforms Ivy Bridge EP, being  roughly 130\,\% faster.  The reason for
that is  two-fold: \emph{i}) it has  higher clock frequency  than Ivy Bridge
(2.6\,GHz  against  2.2\,GHz),  and  \emph{ii})  it  aims at  computing  speed
regardless to  energy consumption.  In our experiments,  Ivy Bridge EP
showed the best energy-to-solution, reducing the energy consumption of
Xeon E5 by approximately 7\,\%.

Averaged TTS and ETS results over 10 repetitions of 24h simulation on \tinto are shown in Figure~\ref{fig:4}. Correponding results for average total power are shown in Table~\ref{tab:3}.  
As expected, the average total power dissipated by \cosmoart while using the blocking MPI policy is about 2\,\% less than that consumed by the polling policy. This is due to the power-friendly (blocking) waits performed by the CPU cores while using the blocking MPI mode, in contrast to the polling waits that take place when the polling MPI mode is set. From the point of view of the execution time, there is a slight reduction for the blocking MPI mode. One can expect that the costs of the interruptions used in the blocking MPI policy can lead to an increase of the total execution time, however, applications with extensive computing, as e.g. \cosmoart, may show better performance with the blocking modes. This effect works in favor with our results since the energy consumed is also reduced (2\,\%) just by changing the behavior of the OpenMPI library to the blocking policy, i.e, with the \texttt{--mca mpi\_yield\_when\_idle} parameter set in the \texttt{mpirun} command.

@@

\begin{figure}[htbf]
  \includegraphics[width=0.5\textwidth]{Figs/Time_E2S_COSMO-ART-5.eps}
  \caption{Mean time-to-solution and energy-to-solution on
    \tinto for a 24h simulation using the polling and blocking MPI policies. The 95\,\% confidence intervals of Student's $t$ distribution (10 samples) clearly indicates 
    an improvement for blocking policy in both TTS and ETS.}
  \label{fig:4}
\end{figure}

@@

A full tracing experiment is conducted on \tinto in order to capture an overall power
profile at a much finer resolution using the attached PDUs in each node. First we run \cosmoart 
for a full-day simulation and analyze the power consumption. Second we correlate performance 
with power traces of two 1-hour time frames during the day: mid-day (11h--12h) and
midnight (23h--24h). All the experiments were performed using 192 \cosmoart processes
m 12 processes in each of node.

% For all the experiments we also analyze the contribution of the MPI library to the energy 
% consumption and how the use of the blocking versus polling message-passing policies can 
% potentially render energy savings. Specifically, the OpenMPI 1.6.5 library installed in \tinto, 
% features two operation modes, blocking and modes, which can be selected before \cosmoart is 
% launched. In the polling configuration (the default mode) ``MPI'' continually polls
% the network interface to check for the completion of an event (e.g. send or receive).
% This usually yields low latency but high CPU utilization. Thus, can one expect that this mode attains the best performance, possibly at the cost a higher energy usage. In the blocking
% mode, the CPU is yield to other processes/threads if there are no incoming messages.

In order to obtain an initial overview of the power dissipation of the system model, we obtained power profiles leveraging our \pmlib power-measurement framework over a full-day simulation of \cosmoart. These profiles, obtained with both polling and blocking MPI policies, are depicted in  Figure~\ref{fig:9}. The first observation for the polling MPI policy is that the first and last hours of the day dissipate slightly less power than the mid-day hours. We relate these variations due to the chemical reactions and aerosols computed by ART, that are taking place during daylight hours but not during night. In this case peak power consumption is about 3821\,W. Looking at the power profile using the blocking MPI policy, we observe that this pattern is repeated, favoring the reduction of power during night hours in a more noticeable way than that for the polling case.The peak power dissipation during day hours in this mode is about 3768\,W, which means a reduction of around 50\,W with respect to the polling mode. Also, we notice systematic power drops in each hour simulated. We relate them to periods in which the processes were blocked waiting for incoming messages or synchronizations, thus yielding the CPU to low usage during. Our final observation is the reduction of the TTS in a factor of 2\,\% with the blocking MPI policy with respect to the polling one.

To gain more insights about the power-performance behavior of \cosmoart, we obtained power-per\-for\-man\-ce traces so that they allow us to correlate the type of tasks being executed with the total power dissipated at each time (see Figure~\ref{fig:11}). For this experiment we limit the simulation time to 1 hour because \emph{i}) the pattern of the computation performed by \cosmoart is repeated hourly and, \emph{ii}) the weight of a full-day simulation trace can not be easily handled using our machines and memory available (each trace file was about 45\,GB). To face the problem we shrink the simulations to 1 hour and test a time frame during midday (from 11h to 12h) and during midnight (from 23h to 24h), in which \cosmoart perform different internal computations due to the different chemical reactions that happen during day and night. At the same time we also leverage the two already stated MPI policies: polling and blocking.

All the traces in Figure~\ref{fig:11} are divided in different phases. In first phase all the processes perform group communications and execute MPI-receive primitives (e.g. \texttt{MPI\_Recv}). The second performs computation bursts, in between group communications, blocking sends, immediate receives and MPI-wait primitives are executed. Finally, group communications interlaced with computation and a significant synchronization phase is performed. From the performance point of view, one can notice a slight reduction of the TTS that blocking MPI mode leads over the polling MPI policy (around 2\,\%). A reduction of power of about  using the blocking MPI policy is produced only in routines that actually perfom synchronizations, group communications and MPI-receive/-wait primitives. Thus, the reduction of the pair power-energy directly depends on the percentage of time in which blocking MPI operations are executed. For our 1 hour simulation traces, an average of 50\,\% of the time the processes were blocked. Taking into account that  the total power reduction during blocking phases is about 5\,\%, one can compute the reduction of power, i.e., $50\,\% \times 5\,\% = 2.5\%$, which is consistent power reduction obtained for the full-day simulation (2\,\%).

\begin{figure*}[ht]
  \centering
  \hspace{0.8cm}
  \scalebox{0.55}{\input{Figs/24hour.pstex_t}}
  \caption{24 hours simulation trace using the MPI blocking mode on \tinto.}
  \label{fig:9}
\end{figure*}

@@

\begin{figure*}[ht]
  \centering
  \scriptsize
  Simulation of 1 hour during midday (11h--12h) leveraging the polling MPI policy\\
  \scalebox{0.5}{\input{Figs/11_12_blq0_figure.pstex_t}}\\
  Simulation of 1 hour during midday (11h--12h) leveraging the blocking MPI policy\\
  \scalebox{0.5}{\input{Figs/11_12_blq1_figure.pstex_t}}\\
  1 hour simulation during midnight (23h--24h) leveraging the polling MPI policy\\
  \scalebox{0.5}{\input{Figs/23_24_blq0_figure.pstex_t}}\\
  1 hour simulation during midnight (23h--24h) leveraging the blocking MPI policy\\
  \scalebox{0.5}{\input{Figs/23_24_blq1_figure.pstex_t}}\\
  \begin{tabular}{rlp{-0.2cm}rlp{-0.2cm}rlp{-0.2cm}rlp{-0.2cm}rl}
& &  & &  & &  \\[-0.15cm]
Computation     & \multicolumn{1}{>{\columncolor[RGB]{  0,  , 255}}p{0.4cm}}{}  & & 
Block. send     & \multicolumn{1}{>{\columncolor[RGB]{255,  0,174}}p{0.4cm}}{}  & & 
Synchronization & \multicolumn{1}{>{\columncolor[RGB]{179,  0,  0}}p{0.4cm}}{}  & & 
Wait/Wait all   & \multicolumn{1}{>{\columncolor[RGB]{235,255,  0}}p{0.4cm}}{}  \\
& &  & &  & &  \\[-0.15cm]
Immediate recv. & \multicolumn{1}{>{\columncolor[RGB]{100,100,177}}p{0.4cm}}{}  & &
Waiting a msg.  & \multicolumn{1}{>{\columncolor[RGB]{255,  0  ,0}}p{0.4cm}}{}  & & 
Group comm.     & \multicolumn{1}{>{\columncolor[RGB]{255,144, 26}}p{0.4cm}}{}  & &
Others          & \multicolumn{1}{>{\columncolor[RGB]{192,224,  0}}p{0.4cm}}{}  \\
%I/O             & \multicolumn{1}{>{\columncolor[RGB]{172,174, 41}}p{0.4cm}}{}  & &
%
%Idle            & \multicolumn{1}{>{\columncolor[RGB]{117,195,255}}p{0.4cm}}{}  \\
\end{tabular}
  \caption{Power-performance traces during midday and midnight leveraging the polling and blocking MPI policies on \tinto.}
  \label{fig:11}
\end{figure*}
